---
title: "Systemfit package"
author: "Olga Mierzwa"
date: "16.05.2015"
output: pdf_document
---

# Introduction to systemfit
Many statistical analyses requires bulding models containing systems of structurally related equations. The systemfit package offers the possibilty to estimate systems of linear and non-linear equations within the R language. Package offers commonly use econemtrics techniques for estimation: “ordinary least squares”
(OLS), “seemingly unrelated regression” (SUR), and the instrumental variable (IV) methods
“two-stage least squares” (2SLS) and “three-stage least squares” (3SLS), where SUR
and 3SLS estimations can optionally be iterated. Moreover, the systemfit package enable statistical testing.

# Systemfit functions
Systemfit package was created by Arne Henningsen and Jeff D. Hamann and is publically availabe on CRAN. The package contains 32 functions. The most basic function is **systemfit** function and this document will focus on introducing it's main functionality.

## Basic systemfit function
The **systemfit** function is used to estimate systems of linear equations using Ordinary Least Squares (OLS), Weighted Least Squares (WLS), Seemingly Unrelated Regression (SUR), Two-Stage Least Squares (2SLS), Weighted TwoStage Least Squares (W2SLS) or Three-Stage Least Squares (3SLS). The systemfit follows lm function functionality in terms of user interface and results. The econometrics formulas used in the systemfit package estimation are described in [systemfit: A Package for Estimating Systems of
Simultaneous Equations in R][1]


### Using systemfit
Systemfit function contains many arguments however there is only one which is always required - **formula**. It can be either a single formula to estimate a single model or a list of formulas to estimate the whole system. Each formula is a standard regression model in R language.

Data for the systemfit are sourced from environment unless speciied by an optional parameter **data** which should be a data frame containing the variables in the model.

```{r, message =FALSE}
library(systemfit)
data( "Kmenta" )
eq1 <- consump ~ price + income
eq2 <- consump ~ price + farmPrice + trend
system <- list( demand = eq1, supply = eq2 )
## estimation (be default OLS)
eqRes <- systemfit( system, data=Kmenta )
```

Commonly used argument is **method**. It is a string object that represents the estimation method: "OLS", "WLS", "SUR", "2SLS", "W2SLS", or "3SLS".

```{r}
## iterated SUR estimation
fitsur <- systemfit( system, "SUR", data = Kmenta )
```

In case a modeller would like to estimate "IV" model, she should include **ins** argument. It should be provided in the following form. One-sided model formula specifying instrumental variables or a list of one-sided model formulas if instruments varies between different equations.

In case modeller wishes to impose restrictions on coefficients in the quations she has can express them by:

* **restrict.matrix** an optional m x n matrix where m equals to the number of restrictions, n is  the number of all coefficients from all equations.
There is a possibility to provide the restrictions in symbolic form.

```{r}
restrict <- "demand_price + supply_price = 0"
fitsurRmat <- systemfit(system, method = "SUR", data = Kmenta, restrict.matrix = restrict)
```

* **restrict.rhs** an optional vector with j elements to impose linear restrictions with
default which cosists of j zeros.

```{r}
Rmat <- matrix(0, nrow = 1, ncol = 7)
Rmat[1, 2] <- 1
Rmat[1, 6] <- 1
qvec <- c(0)
fitsurRmatNum <- systemfit(system, method = "SUR", data = Kmenta, restrict.matrix = Rmat, restrict.rhs = qvec)
```

* **restrict.regMat** an optional matrix to impose restrictions on the coefficients by post-multiplying
the regressor matrix with this matrix.

```{r}
modRegMat <- matrix(0, nrow = 7, ncol = 6)
modRegMat[1:5, 1:5] <- diag(5)
modRegMat[6, 2] <- -1
modRegMat[7, 6] <- 1
fitsurRegMat <- systemfit(system, method = "SUR", data = Kmenta, restrict.regMat = modRegMat)
```

The rest of the possible systemfit parameters are described in [package documentation][2].

### Returned objects
Retured object of systemfit returns is a list of the class systemfit that consists of the all results that are producted by the entire system. One object of this list is "eq" that is also a list and contains one object for each estimated equation. Each "eq" objects are of the special class systemfit.equation and corresponds to the results for specific equation.

```{r}
fitols <- systemfit( system, data=Kmenta )
print( fitols )
```

The **summary** method can be used to compute and print summary results of systemfit.
```{r}
summary(fitols)
```

In order to reduce the amount of the printed outcome use the optional arguments residCov and equations.
Argument **residCov** controls printing the covariance matrix and the correlation
matrix of the residuals. Argument **equations** controls the equations results.
```{r}
summary(fitols, residCov = FALSE, equations = FALSE)
```

## Testing linear restrictions

Systemfit allows to test linear restrictions by an F test, two Wald tests and
an LR test. Testing can be done using method **linearHypothesis** for the F test and
Wald tests, **lrtest** method for LR tests.

### Method linearHypothesis

Linear.hypothesis needs unrestricted regression from systemfit as it's first argument. The second and third arguments are the restriction matrix and the optional vector. The optional argument test is a character string, "FT", "F", or "Chisq", specifying the test to conduct.

```{r}
linearHypothesis(fitsur, Rmat, qvec, test = "FT")

linearHypothesis(fitsur, Rmat, qvec, test = "F")

linearHypothesis(fitsur, Rmat, qvec, test = "Chisq")
```

### Method lrtest
All arguments of the lrtest method for systemfit should be fitted model objects from systemfit. 

```{r}
lrtest(fitsurRmat, fitsur)
```

### Tests comparison
All tests check the same hypothesis and suggest the same decision: The
null hypothesis cannot be rejected at any reasonable level of significance.


| Tests         | Results (Pr)  |
| ------------- |:-------------:|
| FT            | 0.3413        |
| F             | 0.4407        |
| Chisq         | 0.4351        |
| LR            | 0.3899        |

[1]: http://www.jstatsoft.org/v23/i04/paper
[2]: http://cran.r-project.org/web/packages/systemfit/systemfit.pdf